### 2015-06-11 ApTranscriptome_TR.Rmd
in response to reviewer 2, changed comparisons between species to focus on only the transcripts with a significant species x temperature interaction. previously, this also included transcripts with a main effect of temperature only.

---
output: word_document
---
Thermal reactionome of the common ant species *Aphaenogaster picea* and *A. carolinensis*
========================================================================================

@@

**October 14, 2014**

@@

```{r setup, echo=FALSE, results='hide', message = FALSE}
# Global settings
options(stringsAsFactors=FALSE, digits = 3, scipen = 1)

@@

# Load Bioconductor libraries
source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)	
library(GOSemSim)

@@

In this technical report, which accompanies the manuscript **Patterns of stress tolerance versus resistance in genome-wide expression data of parapatric ant species** (Stanton-Geddes et al., in press), we:

1. describe the *de novo* assembly of the transcriptome for two ant species within the *Aphaenogaster rudis-picea-fulva* species complex `r citep("10.1155/2012/752815")`
2. identify thermally-responsive genes
3. evaluate differences in reaction norms of expression for each thermally-responsive gene between the two species
4. perform gene set enrichment analysis of thermally-responsive genes for the two species

This script is completely reproducible assuming that R, `knitr` and the other required libraries (listed within the source document) are installed on a standard linux system using the following:

    Rscript -e "library(knitr); knit('ApTranscriptome_TR.Rmd')"

@@

# merge results
Ap.response.type <- rbind(interaction.response.type, temperature.response.type)
colnames(Ap.response.type)[which(colnames(Ap.response.type) == ".id")] <- "Transcript"
Ap.response.type <- data.table(Ap.response.type)
setkey(Ap.response.type, Transcript)
str(Ap.response.type)

# save results to file
write.table(file = paste(resultsdir, "Ap_responsive_transcripts_", Sys.Date(), ".txt", sep = ""), Ap.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

```{r temperature_response}
A22.type.table <- table(Ap.response.type$A22.type)
Ar.type.table <- table(Ap.response.type$Ar.type)

@@

```{r mosaic_plot}
# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Acar = Ap.response.type$Ar.type, Apic = Ap.response.type$A22.type)
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

```{r ed_cartoon}
enhanced_defense_dat <- md_sub1
enhanced_defense_dat$value <- 0
# set non-shared response to random negative value 
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Low"), "value"] <- runif(5, min = -300, max = 0)
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Bimodal"), "value"] <- runif(5, min = -300, max = 0)
# change shared response to positive
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Low" & enhanced_defense_dat$Apic == "Low"), "value"] <- 500
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Bimodal" & enhanced_defense_dat$Apic == "Bimodal"), "value"] <- 300

@@

```{r tol_cartoon}
tolerance_dat <- md_sub1
tolerance_dat$value <- 0
# set non-shared response to random positive value 
tolerance_dat[which(tolerance_dat$Acar == "Low"), "value"] <- runif(5, min = 100, max = 300)
tolerance_dat[which(tolerance_dat$Acar == "Bimodal"), "value"] <- runif(5, min = 100, max = 200)
# change shared response to near zero
tolerance_dat[which(tolerance_dat$Acar == "Low" & tolerance_dat$Apic == "Low"), "value"] <- 30
tolerance_dat[which(tolerance_dat$Acar == "Bimodal" & tolerance_dat$Apic == "Bimodal"), "value"] <- 20

@@

```{r gen_assimilation_high}
Ap.high.Ac.nr <- Ap.response.type[which(Ap.response.type$A22.type == "High" & Ap.response.type$Ar.type == "NotResp"), ]

@@

```{r gen_assimilation_bim}
Ap.bim.Ac.nr <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "NotResp"), ]

@@

```{r max_exp_PDF, warning = FALSE, message = FALSE}
# reshape data
Ap.df <- data.frame(Transcript = rep(Ap.response.type$Transcript, times = 2), colony = rep(c("ApVT", "AcNC"), each = length(Ap.response.type$Transcript)), max.val = c(Ap.response.type$A22.max, Ap.response.type$Ar.max))

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts.df <- Ap.response.type[which(Ap.response.type$A22.type == "High"), ]

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts.df <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), ]

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" | Ap.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

# merge optimum expression, in Ap.response.type data.frame, with predicted TPM data
Ap.induction.wide <- data.table(Ap.induction.wide)
setkey(Ap.induction.wide, Transcript)
Ap.response.type2 <- Ap.induction.wide[Ap.response.type]
```

@@

```{r log_inducibility}
Ap.response.type2.A22.high <- Ap.response.type2[Ap.response.type2$Transcript %in% A22.high.transcripts.df$Transcript, ]
str(Ap.response.type2.A22.high)

# filter out NotResp transcripts in Ar
Ap.response.type2.A22.high <- Ap.response.type2.A22.high[which(Ap.response.type2.A22.high$Ar.type != "NotResp"), ]

# ratio of upregulation in A22 vs Ar
Ap.response.type2.A22.high$A22.Ar.upreg <- Ap.response.type2.A22.high$A22.upreg / Ap.response.type2.A22.high$Ar.upreg

# ratio of constitutive expression in Ar vs A22
Ap.response.type2.A22.high$A22.Ar.constitutive <- Ap.response.type2.A22.high$A22.opt / Ap.response.type2.A22.high$Ar.opt

# plot!
gg11 <- ggplot(Ap.response.type2.A22.high, aes(x = log(A22.Ar.upreg), y = log(A22.Ar.constitutive))) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(x = expression(paste("log(inducibility ", italic("A. picea / A. carolinensis"), ")")),
       y = expression(paste("log(constitutive expression ", italic("A. picea / A. carolinensis"), ")")))

@@

A22.high.lm <- lm(log(A22.Ar.constitutive) ~ log(A22.Ar.upreg), data = Ap.response.type2.A22.high)

@@

```{r assimilation_Ar_low}
### filter for A. carolinensis *Low* transcripts
Ap.response.type2.Ar.low <- Ap.response.type2[Ap.response.type2$Transcript %in% Ar.low.transcripts.df$Transcript, ]
str(Ap.response.type2.Ar.low)

# filter out NotResp transcripts in A22
Ap.response.type2.Ar.low <- Ap.response.type2.Ar.low[which(Ap.response.type2.Ar.low$A22.type != "NotResp"), ]

# ratio of upregulation
Ap.response.type2.Ar.low$Ar.A22.upreg <- Ap.response.type2.Ar.low$Ar.upreg / Ap.response.type2.Ar.low$A22.upreg

# ratio of constitutive expression
Ap.response.type2.Ar.low$Ar.A22.constitutive <- Ap.response.type2.Ar.low$Ar.opt / Ap.response.type2.Ar.low$A22.opt

gg12 <- ggplot(Ap.response.type2.Ar.low, aes(x = log(Ar.A22.upreg), y = log(Ar.A22.constitutive))) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(x = expression(paste("log(inducibility ", italic("A. carolinensis / A. picea"), ")")),
       y = expression(paste("log(constitutive expression ", italic("A. carolinensis / A. picea"), ")")))

@@

Ar.low.lm <- lm(log(Ar.A22.constitutive) ~ log(Ar.A22.upreg), data = Ap.response.type2.Ar.low)

@@

```{r extract_Transcripts}
Ap.response.type <- as.data.frame(Ap.response.type)
A22.high.transcripts <-  Ap.response.type[which(Ap.response.type$A22.type == "High"), "Transcript"]
Ar.high.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "High"), "Transcript"]

A22.low.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Low"), "Transcript"]
Ar.low.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Low"), "Transcript"]

A22.bim.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int.transcripts <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int.transcripts <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r mean_exp}
t.test(log(Ap.response.type$A22.opt), log(Ap.response.type$Ar.opt))
# expression level at optimum temp greater in Ar than A22

# non-parametric test
w4 = wilcox.test(Ap.response.type$A22.opt, 
                 Ap.response.type$Ar.opt,
                 alternative = "two.sided", paired = TRUE, conf.int = TRUE)

@@

```{r ann_responsive, echo=TRUE}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, Ap.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = paste(resultsdir, "Ap_responsive_genes_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

@@

```{r gsea_Ap_high, results = 'hide'}
## Genes with *High* expression in both colonies
Ap.high <- Ap.response.type[which(Ap.response.type$Ar.type == "High" & Ap.response.type$A22.type == "High"), "Transcript"]
Ap.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.high) <- names(Ap.geneList)
Ap.geneList.high[(which(names(Ap.geneList.high) %in% Ap.high))] <- 1
# check correct number of values set to 1
table(Ap.geneList.high)
# run GSEA
Ap.high.gsea <- gsea(genelist = Ap.geneList.high, geneID2GO = geneID2GO, plotpath = NA)
```

@@

```{r gsea_Ap_low, results = "hide"}
## Genes with *Low* expression in both colonies
Ap.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Low" & Ap.response.type$A22.type == "Low"), "Transcript"]
Ap.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.low) <- names(Ap.geneList)
Ap.geneList.low[(which(names(Ap.geneList.low) %in% Ap.low))] <- 1
# check correct number of values set to 1
table(Ap.geneList.low)
# Run GSEA
Ap.low.gsea <- gsea(genelist = Ap.geneList.low, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ap_bim, resuts = 'hide'}
# Genes with *Bimodal* expression in both colonies
Ap.bim <- Ap.response.type[which(Ap.response.type$A22.type == "Bimodal" & Ap.response.type$Ar.type == "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ap.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.bim) <- names(Ap.geneList)
Ap.geneList.bim[(which(names(Ap.geneList.bim) %in% Ap.bim))] <- 1
# check correct number of values set to 1
table(Ap.geneList.bim)
# Run GSEA
Ap.bim.gsea <- gsea(genelist = Ap.geneList.bim, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ap_int, results = 'hide'}
# genes with *Intermediate* expression in both colonies
Ap.int <- Ap.response.type[which(Ap.response.type$A22.type == "Intermediate" & Ap.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)
# run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ar_int_A22_bim, results = 'hide'}
# genes with *Intermediate* expression in Ar and *Bimodal* in A22
Ar.int.A22.bim <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" & Ap.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.int.A22.bim.geneList <- rep(0, length = length(Ap.geneList))
names(Ar.int.A22.bim.geneList) <- names(Ap.geneList)
Ar.int.A22.bim.geneList[(which(names(Ar.int.A22.bim.geneList) %in% Ar.int.A22.bim))] <- 1
# check correct number of values set to 1
table(Ar.int.A22.bim.geneList)
# run GSEA
Ar.int.A22.bim.gsea <- gsea(genelist = Ar.int.A22.bim.geneList, geneID2GO = geneID2GO, plotpath = NA)
```

@@

```{r gsea_Ar_int_A22_low, results = 'hide'}
# genes with *Intermediate* expression in Ar and *Low* in A22
Ar.int.A22.low <- Ap.response.type[which(Ap.response.type$Ar.type == "Intermediate" & Ap.response.type$A22.type == "Low"), "Transcript"]
Ar.int.A22.low.geneList <- rep(0, length = length(Ap.geneList))
names(Ar.int.A22.low.geneList) <- names(Ap.geneList)
Ar.int.A22.low.geneList[(which(names(Ar.int.A22.low.geneList) %in% Ar.int.A22.low))] <- 1
# check correct number of values set to 1
table(Ar.int.A22.low.geneList)
# run GSEA
Ar.int.A22.low.gsea <- gsea(genelist = Ar.int.A22.low.geneList, geneID2GO = geneID2GO, plotpath = NA)
```

>>>>

---
output: pdf_document
---

Technical Report for **Thermal reactionomes reveal divergent responses to thermal extremes in warm and cool-climate ant species**
========================================================================================

@@

**`r format(Sys.time(), '%d %B, %Y')`**

@@

```{r setup, echo=FALSE, results='hide', message = FALSE, warning=FALSE}
# Global settings
options(stringsAsFactors=FALSE, digits = 3, scipen = 1)

@@

# Load Bioconductor libraries
#source("http://bioconductor.org/biocLite.R")
#biocLite("topGO")
#biocLite("Rgraphviz")
#biocLite("GOSemSim")
library(topGO) 
library(Rgraphviz)	
library(GOSemSim)

@@

In this technical report, which accompanies the manuscript **Thermal reactionomes reveal divergent responses to thermal extremes in warm and cool-climate ant species** (Stanton-Geddes et al., in press), we:

1. Describe the *de novo* assembly of the transcriptome for two ant species within the *Aphaenogaster rudis-picea-fulva* species complex `r citep("10.1155/2012/752815")`.
2. Identify thermally-responsive genes.
3. Evaluate differences in transcriptome-wide patterns of thermal responsiveness (i.e. the thermal reactionome) between the two species
4. Perform gene set enrichment analysis of thermally-responsive genes for the two species.

@@

# for subsequent analyses, focus on the transcripts with significant temperature x species interaction only
colnames(interaction.response.type)[which(colnames(interaction.response.type) == ".id")] <- "Transcript"
interaction.response.type <- data.table(interaction.response.type)
setkey(interaction.response.type, Transcript)
str(interaction.response.type)

# save results to file
write.table(file = paste(resultsdir, "Ap_responsive_transcripts_", Sys.Date(), ".txt", sep = ""), interaction.response.type, row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

@@

```{r temperature_response}
A22.type.table <- table(interaction.response.type$A22.type)
Ar.type.table <- table(interaction.response.type$Ar.type)

@@

```{r mosaic_plot}
# Reorganize for plotting to show overlap among categories between colonies
type.table <- table(Acar = interaction.response.type$Ar.type, Apic = interaction.response.type$A22.type)
# reorder
tt2 <- type.table[c("Low", "Intermediate", "High", "Bimodal", "NotResp"), c("Low", "Intermediate", "High", "Bimodal", "NotResp")]

@@

```{r ed_cartoon}
enhanced_defense_dat <- md_sub1
enhanced_defense_dat$value <- 0
# set non-shared response to random negative value 
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Low"), "value"] <- runif(5, min=140, max=150)
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Bimodal"), "value"] <- rep(-200, 5)
# change shared response to positive
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Low" & enhanced_defense_dat$Apic == "Low"), "value"] <- 50
enhanced_defense_dat[which(enhanced_defense_dat$Acar == "Bimodal" & enhanced_defense_dat$Apic == "Bimodal"), "value"] <- 50

@@

```{r tol_cartoon}
tolerance_dat <- md_sub1
tolerance_dat$value <- 0
# set non-shared response to random positive value 
tolerance_dat[which(tolerance_dat$Acar == "Low"), "value"] <- runif(5, min=140, max=160)
tolerance_dat[which(tolerance_dat$Acar == "Bimodal"), "value"] <- runif(5, min=140, max=150)
# change shared response to near zero
tolerance_dat[which(tolerance_dat$Acar == "Low" & tolerance_dat$Apic == "Low"), "value"] <- 20
tolerance_dat[which(tolerance_dat$Acar == "Bimodal" & tolerance_dat$Apic == "Bimodal"), "value"] <- 20

@@

```{r gen_assimilation_high}
Ap.high.Ac.nr <- interaction.response.type[which(interaction.response.type$A22.type == "High" & interaction.response.type$Ar.type == "NotResp"), ]

@@

```{r gen_assimilation_bim}
Ap.bim.Ac.nr <- interaction.response.type[which(interaction.response.type$A22.type == "Bimodal" & interaction.response.type$Ar.type == "NotResp"), ]

@@

```{r max_exp_PDF, warning = FALSE, message = FALSE}
# reshape data
Ap.df <- data.frame(Transcript = rep(interaction.response.type$Transcript, times = 2), colony = rep(c("ApVT", "AcNC"), each = length(interaction.response.type$Transcript)), max.val = c(interaction.response.type$A22.max, interaction.response.type$Ar.max))

@@

```{r optimum_expression_comparison}
# list of transcripts that are 'high' expressed in A22
A22.high.transcripts.df <- interaction.response.type[which(interaction.response.type$A22.type == "High"), ]

@@

```{r Ar_low_wilcoxon}
# list of transcripts that are 'high' expressed in Ar
Ar.low.transcripts.df <- interaction.response.type[which(interaction.response.type$Ar.type == "Low"), ]

@@

```{r intermediate_transcripts}
# list of transcripts that are 'Intermediate' expressed in either colony
Ap.int.transcripts <- interaction.response.type[which(interaction.response.type$Ar.type == "Intermediate" | interaction.response.type$A22.type == "Intermediate"), ]
# T test
t.test(log(Ap.int.transcripts$A22.opt+1), log(Ap.int.transcripts$Ar.opt+1))
# Wilcoxon signed rank-test
w3 <- wilcox.test(Ap.int.transcripts$A22.opt, Ap.int.transcripts$Ar.opt, alternative = "two.sided", paired = TRUE, conf.int = TRUE)
w3
```

@@

# merge optimum expression, in interaction.response.type data.frame, with predicted TPM data
Ap.induction.wide <- data.table(Ap.induction.wide)
setkey(Ap.induction.wide, Transcript)
interaction.response.type2 <- Ap.induction.wide[interaction.response.type]
```

@@

```{r log_inducibility}
interaction.response.type2.A22.high <- interaction.response.type2[interaction.response.type2$Transcript %in% A22.high.transcripts.df$Transcript, ]
str(interaction.response.type2.A22.high)

# filter out NotResp transcripts in Ar
interaction.response.type2.A22.high <- interaction.response.type2.A22.high[which(interaction.response.type2.A22.high$Ar.type != "NotResp"), ]

# ratio of upregulation in A22 vs Ar
interaction.response.type2.A22.high$A22.Ar.upreg <- interaction.response.type2.A22.high$A22.upreg / interaction.response.type2.A22.high$Ar.upreg

# ratio of constitutive expression in Ar vs A22
interaction.response.type2.A22.high$A22.Ar.constitutive <- interaction.response.type2.A22.high$A22.opt / interaction.response.type2.A22.high$Ar.opt

# plot!
gg11 <- ggplot(interaction.response.type2.A22.high, aes(x = log(A22.Ar.upreg), y = log(A22.Ar.constitutive))) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(x = expression(paste("log(inducibility ", italic("A. picea / A. carolinensis"), ")")),
       y = expression(paste("log(constitutive expression ", italic("A. picea / A. carolinensis"), ")")))

@@

A22.high.lm <- lm(log(A22.Ar.constitutive) ~ log(A22.Ar.upreg), data = interaction.response.type2.A22.high)

@@

```{r assimilation_Ar_low}
### filter for A. carolinensis *Low* transcripts
interaction.response.type2.Ar.low <- interaction.response.type2[interaction.response.type2$Transcript %in% Ar.low.transcripts.df$Transcript, ]
str(interaction.response.type2.Ar.low)

# filter out NotResp transcripts in A22
interaction.response.type2.Ar.low <- interaction.response.type2.Ar.low[which(interaction.response.type2.Ar.low$A22.type != "NotResp"), ]

# ratio of upregulation
interaction.response.type2.Ar.low$Ar.A22.upreg <- interaction.response.type2.Ar.low$Ar.upreg / interaction.response.type2.Ar.low$A22.upreg

# ratio of constitutive expression
interaction.response.type2.Ar.low$Ar.A22.constitutive <- interaction.response.type2.Ar.low$Ar.opt / interaction.response.type2.Ar.low$A22.opt

gg12 <- ggplot(interaction.response.type2.Ar.low, aes(x = log(Ar.A22.upreg), y = log(Ar.A22.constitutive))) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  labs(x = expression(paste("log(inducibility ", italic("A. carolinensis / A. picea"), ")")),
       y = expression(paste("log(constitutive expression ", italic("A. carolinensis / A. picea"), ")")))

@@

Ar.low.lm <- lm(log(Ar.A22.constitutive) ~ log(Ar.A22.upreg), data = interaction.response.type2.Ar.low)

@@

```{r extract_Transcripts}
interaction.response.type <- as.data.frame(interaction.response.type)
A22.high.transcripts <-  interaction.response.type[which(interaction.response.type$A22.type == "High"), "Transcript"]
Ar.high.transcripts <- interaction.response.type[which(interaction.response.type$Ar.type == "High"), "Transcript"]

A22.low.transcripts <- interaction.response.type[which(interaction.response.type$A22.type == "Low"), "Transcript"]
Ar.low.transcripts <- interaction.response.type[which(interaction.response.type$Ar.type == "Low"), "Transcript"]

A22.bim.transcripts <- interaction.response.type[which(interaction.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.bim.transcripts <- interaction.response.type[which(interaction.response.type$Ar.type == "Bimodal"), "Transcript"]

A22.int.transcripts <- interaction.response.type[which(interaction.response.type$A22.type == "Intermediate"), "Transcript"]
Ar.int.transcripts <- interaction.response.type[which(interaction.response.type$Ar.type == "Intermediate"), "Transcript"]
```

@@

```{r mean_exp}
t.test(log(interaction.response.type$A22.opt), log(interaction.response.type$Ar.opt))
# expression level at optimum temp greater in Ar than A22

# non-parametric test
w4 = wilcox.test(interaction.response.type$A22.opt, 
                 interaction.response.type$Ar.opt,
                 alternative = "two.sided", paired = TRUE, conf.int = TRUE)

@@

```{r ann_responsive, echo=TRUE}
responsive.lms.ann.type <- signif.transcripts[names(responsive.lms)]
# combine responsive.lms.ann with "type" information
responsive.lms.ann.type <- merge(responsive.lms.ann.type, interaction.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.lms.ann.type)
write.csv(responsive.lms.ann.type, file = paste(resultsdir, "Ap_responsive_genes_", Sys.Date(), ".csv", sep = ""), row.names = FALSE)
```

@@

```{r gsea_Ap_high, results = 'hide'}
## Genes with *High* expression in both colonies
Ap.high <- interaction.response.type[which(interaction.response.type$Ar.type == "High" & interaction.response.type$A22.type == "High"), "Transcript"]
Ap.geneList.high <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.high) <- names(Ap.geneList)
Ap.geneList.high[(which(names(Ap.geneList.high) %in% Ap.high))] <- 1
# check correct number of values set to 1
table(Ap.geneList.high)
# run GSEA
Ap.high.gsea <- gsea(genelist = Ap.geneList.high, geneID2GO = geneID2GO, plotpath = NA)
```

@@

```{r gsea_Ap_low, results = "hide"}
## Genes with *Low* expression in both colonies
Ap.low <- interaction.response.type[which(interaction.response.type$Ar.type == "Low" & interaction.response.type$A22.type == "Low"), "Transcript"]
Ap.geneList.low <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.low) <- names(Ap.geneList)
Ap.geneList.low[(which(names(Ap.geneList.low) %in% Ap.low))] <- 1
# check correct number of values set to 1
table(Ap.geneList.low)
# Run GSEA
Ap.low.gsea <- gsea(genelist = Ap.geneList.low, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ap_bim, resuts = 'hide'}
# Genes with *Bimodal* expression in both colonies
Ap.bim <- interaction.response.type[which(interaction.response.type$A22.type == "Bimodal" & interaction.response.type$Ar.type == "Bimodal"), "Transcript"]
# create gene list, setting value to 1 for "bim" transcripts
Ap.geneList.bim <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.bim) <- names(Ap.geneList)
Ap.geneList.bim[(which(names(Ap.geneList.bim) %in% Ap.bim))] <- 1
# check correct number of values set to 1
table(Ap.geneList.bim)
# Run GSEA
Ap.bim.gsea <- gsea(genelist = Ap.geneList.bim, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ap_int, results = 'hide'}
# genes with *Intermediate* expression in both colonies
Ap.int <- interaction.response.type[which(interaction.response.type$A22.type == "Intermediate" & interaction.response.type$Ar.type == "Intermediate"), "Transcript"]
# create gene list, setting value to 1 for "int" transcripts
Ap.geneList.int <- rep(0, length = length(Ap.geneList))
names(Ap.geneList.int) <- names(Ap.geneList)
Ap.geneList.int[(which(names(Ap.geneList.int) %in% Ap.int))] <- 1
# check correct number of values set to 1
table(Ap.geneList.int)
# run GSEA
Ap.int.gsea <- gsea(genelist = Ap.geneList.int, geneID2GO = geneID2GO)
```

@@

```{r gsea_Ar_int_A22_bim, results = 'hide'}
# genes with *Intermediate* expression in Ar and *Bimodal* in A22
Ar.int.A22.bim <- interaction.response.type[which(interaction.response.type$Ar.type == "Intermediate" & interaction.response.type$A22.type == "Bimodal"), "Transcript"]
Ar.int.A22.bim.geneList <- rep(0, length = length(Ap.geneList))
names(Ar.int.A22.bim.geneList) <- names(Ap.geneList)
Ar.int.A22.bim.geneList[(which(names(Ar.int.A22.bim.geneList) %in% Ar.int.A22.bim))] <- 1
# check correct number of values set to 1
table(Ar.int.A22.bim.geneList)
# run GSEA
Ar.int.A22.bim.gsea <- gsea(genelist = Ar.int.A22.bim.geneList, geneID2GO = geneID2GO, plotpath = NA)
```

@@

```{r gsea_Ar_int_A22_low, results = 'hide'}
# genes with *Intermediate* expression in Ar and *Low* in A22
Ar.int.A22.low <- interaction.response.type[which(interaction.response.type$Ar.type == "Intermediate" & interaction.response.type$A22.type == "Low"), "Transcript"]
Ar.int.A22.low.geneList <- rep(0, length = length(Ap.geneList))
names(Ar.int.A22.low.geneList) <- names(Ap.geneList)
Ar.int.A22.low.geneList[(which(names(Ar.int.A22.low.geneList) %in% Ar.int.A22.low))] <- 1
# check correct number of values set to 1
table(Ar.int.A22.low.geneList)
# run GSEA
Ar.int.A22.low.gsea <- gsea(genelist = Ar.int.A22.low.geneList, geneID2GO = geneID2GO, plotpath = NA)
```
